\documentclass[aps,rmp,twocolumn,amsmath,amssymb,nofootinbib,superscriptaddress]{revtex4}

\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\op}[2]{\hat{\textbf{#1}}_{#2}}
\newcommand{\dagop}[2]{\hat{\textbf{#1}}_{#2}^\dag}
\usepackage[pdftex]{graphicx}
\usepackage{mathrsfs}
\usepackage[colorlinks]{hyperref}

\begin{document}

\bibliographystyle{apsrev}

%
% Title
%

\title{Competitive information dynamics \& the theory of competitive traits}

%
% Authors
%

\author{Peter P. Rohde}
\email[]{dr.rohde@gmail.com}
\homepage{http://www.peterrohde.org}
\affiliation{Centre for Quantum Software \& Information (UTS:QSI), University of Technology Sydney, Australia}

\date{\today}

\frenchspacing

%
% Abstract
%

\begin{abstract}
\end{abstract}

\maketitle

\tableofcontents

\section{Definitions}

Strategies are associated with a trait, not with the group itself.

Trait abstractly carries algorithmic information content, not attached to a particular physical medium.

Strategy is defined as being between a trait group and its compliment. Gives rise to indirect competition.

Trait competitiveness, $C_t$: score for extent to which a group member $g\in G_t$ abiding by the trait group's strategy $S_{G_t}$ enhances the survival rate of trait $t$.

Discuss competing vs non-competing traits. Local strategy optimisation globally optimises.

There is competition between non-complimentary trait group, i.e partially overlapping. Explains cooperation between groups with overlapping interests.

Traits are ratio of expression, not just binary

An individual's tendency to obey a strategy: $T_{g,S}$

Inter-trait enhancement. $t_1$ ehnahnces expression of $t_2$ - cooperation

Indirect competition: piggbacking off the strategy employed by a correlated trait.

If traits are perfectly correlated, their respective trait groups are necessarily identical, $G_{t_1}=G_{t_2}$. Nonetheless, the strategies associated with the traits may be distinct, $\mathcal{S}(t_1)\neq\mathcal{S}(t_2)$.

Associative traits and trait substitution. Adopting traits.

Collective strategy when S(g1 union g2) neq S(g1) union S(g2))

self-competition for strategically comaptible goals:
for distinct strategies for different traits, where the traits are highly correlated or subset, better strategy is for the weaker trait to piggyback off the stronger common trait than to ouruse its own. strong supergroup strategy leads to weaker subgroup strategy to yield pririty to the supergroup strategy.

trait absirbtion by strategies advancing correlated traits. 

neutral strategies for non-competing traits. eg if a trait is random, with no capacity to influence its own future propagation, we expect a neutral strategy to be observed. i.e the trait is non-competitive.

\subsection{Traits}

A \textit{trait} is an entirely arbitrary characteristic that may be associated with an individual. It may be genotypic, phenotypic, memetic, cultural, hereditary or non-hereditary, acquired or inate, physical or psychological -- it may be defined arbitrarily and abstractly as any feature by which individuals within a group can be identified as having in common.

We use $t$ to denote a single trait, or $\vec t=\{t_1,\dots,t_n\}$ to denote a set of traits.

\subsubsection{Correlated traits}

\subsubsection{Competitive vs non-competitive traits}

Selective pressures. Take two perfectly correlated traits. One may act competitively, the other not. The later inherits all benefits of the former, without actively pursuing a coordinated strategy of its own. Indirectly competitive. To what extent 

\subsection{Individuals \& groups}

We use $U$ to denote the \textit{universe} -- the set of all individuals participating in the model under consideration. An \textit{individual} is a single participant within the universe, characterised entirely by the complete set of traits it exhibits,
\begin{align}
	g &= \{t_1,\dots,t_n\},\nonumber\\
	g &\in U.
\end{align}

A \textit{group} is a set of individuals,
\begin{align}
	G = \{g_1,\dots,g_n\}.	
\end{align}

In the case of \textit{acquired traits}, an individual's trait set is dynamic, undergoing the update rule,
\begin{align}
	g\to g\cup t,	
\end{align}
where $g$ acquires trait $t$.

A \textit{trait group} is the set of all individuals exhibiting a trait or set of traits,
\begin{align}
	G_t = \bigcup_{i\in U} g_i \,|\, t\in g_i.
\end{align}
The complimentary \textit{non-trait group} is denoted $\bar G_t$. Together they satisfy the relations,
\begin{align}
G_t \cup \bar G_t &= U \,\,\text{(completeness)},\nonumber\\
G_t \cap \bar G_t &= \emptyset \,\,\text{(non-overlapping)},
\end{align}
since traits are binary.

\subsection{Strategies}

A \textit{strategy} is defined in a game-theoretic context as the algorithm employed by one player when interacting with another. A player may be an individual or a group of individuals. Thus, strategies may be defined at an individual level (\textit{individual strategies}), or collectively at a group level (\textit{group strategies}),
\begin{align}
	\mathcal{S}(p|q).	
\end{align}

%The joint strategy employed by $n$ players $p_i$ against $m$ players $q_j$ is denoted,
%\begin{align}
%	\mathcal{S}(p_1,\dots,p_n|q_1,\dots,q_m).
%\end{align}
%The special case of $n=1$ refers to a single player engaging in a strategy independently of other players (i.e not a joint strategy).

We adopt the notation $\mathcal{S}'$ for the opposing strategy,
\begin{align}
	\mathcal{S}(p|q) = \mathcal{S}'(q|p).
\end{align}
Two opposing strategies may be chosen entirely independently -- they could be entirely different, or identical.

Individuals within a group may not behave homogenously. Thus, the strategy employed by individuals within a group may in general be distinct from their collective group strategy, which described by the cumulative action of all participating individuals. Thus, in general,
\begin{align}
	\mathcal{S}(g\in G|q)\neq \mathcal{S}(G|q).
\end{align}

\subsubsection{Indirect strategies}

A trait might directly pursue its own competitive strategy, or indirectly inherit it via correlated traits doing so. For example, a genetically induced trait might be highly competitive, but the responsible genes additionally give rise to other secondary traits that are not inherently competitive in their own right. In this instance, the secondary correlated traits indirectly inherit competitiveness from the more competitive primary trait.

\subsection{Payoff}

The \textit{payoff} of a strategy for player $p$, $\mathcal{P}_p(\mathcal{S})$, is the expectation of `reward' (in arbitrary units of utility) obtained by executing a strategy. The payoff may be positive (for `successful' strategies advancing the player's interests), or negative (for `unsuccessful' ones).

For a trait $t$ to develop a collective strategies to guarantee its propagation, their objective is to maximise the joint payoff of the respective trait group, $G_t$.

A special case arises when players compete for finite resources, in which case the sum of their respective payoffs is bounded by the resources' abundancy $c$,
\begin{align}
	\mathcal{P}_p(\mathcal{S}_p) + \mathcal{P}_q(\mathcal{S}_q) \leq c.
\end{align}

Another specialised case is \textit{zero-sum games} in which players' payoffs are exactly opposite, i.e gains made by one come at the direct expense of the other,
\begin{align}
\mathcal{P}_p(\mathcal{S}_p) = -\mathcal{P}_q(\mathcal{S}_q).
\end{align}

Alternately, opposing strategies may collectively act in one another's mutual benefit -- a symbiotic strategy,
\begin{align}
\mathcal{P}_p(\mathcal{S}_{p\cup q}) > \mathcal{P}_p(\mathcal{S}_p).
\end{align}
Here the two distinct players effectively engage in a cooperative inter-group strategy.

\subsubsection{Individual payoff}

\subsubsection{Group payoff}

The payoff for a group is simply given by the sum of the payoffs benefitting the group by all individuals in the universe,
\begin{align}
	\mathcal{P}_{G}(\mathcal{S}(G)) = \sum_{g\in U} \mathcal{P}_G(g)
\end{align}

In the case of a non-competing trait that achieves payoff indirectly via its correlation with competing traits, we can observe properties of the form,
\begin{align}
	\mathcal{P}_{G_1}(\emptyset) \propto \mathcal{P}_{G_2}(\mathcal{S})
\end{align}
That is, the group $G_1$ benefits passively (implementing no strategy of its own) via the payoff bestowed upon group $G_2$ pursuing its own strategy.

\section{Selfish vs altruistic behaviour}

Conditions where pursuing self-interest is congruent with group interest.

\section{Meta-strategy}

Strategic traits: the trait of employing a strategy, $t=\mathcal{S}$. Payoff for strategic trait group is net survival of strategy itself, not a particular group employing the strategy. Strategy is aiming to advance itself, using groups employing it as hosts for itself. May not care about a specific group, and willing to sacrifice them so long as the strategy is preserved long-term. In the case of zero-sum games, might act indifferently.

Notion of meta-strategy: actual goal of a strategy is to advance the strategy group, not an individual group employing the strategy. When the strategy spreads across multiple groups which evolve independently, it undergoes evolution at the meta-level. Now the strategy is purely informational, and is no longer limited to a particular host residing in a particular medium. It can switch across mediums and evolve as such. If groups employing old mediums die out, the strategy survives via implementation in the new medium.

\section{Quantum information -- A new era for strategic dominance}

Future quantum strategies. Transition in computational complexity of strategies.

\section{Humanity}

Philosophical: for humanity, our group strategies are successful at promoting the strategy. Unsuccessful at promoting the groups. We must realign our strategies to become congruent with the trait group `humanity'. Enhance human interests, rather than the strategy. This may require making our old strategies extinct, and artificially replacing them with a new meta-group strategy. We can do this using our capacity for intelligence, computation, reason, and morality.

Identity politics.

\end{document}
